---
title: "EDA 10/2"
author: "Kennedy Wade"
date: "2025-10-02"
output: github_document
---

loading datasets 

```{r}
library(p8105.datasets)
library(dplyr)
library(lubridate)
library(ggplot2)
library(tidyverse)
```

loading the data 

```{r}
data("weather_df")

weather_df = 
  weather_df |>
  mutate(
    month = lubridate::floor_date(date, unit = "month"))

```

Make plots

```{r}
weather_df |> 
  ggplot(aes(x = prcp)) + 
  geom_histogram()
  
```

check extreme values 

```{r}
weather_df |> 
  filter(prcp >= 1000)
```

look at the data again

```{r}
weather_df |> 
  filter(tmax >= 20, tmax <= 30) |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) + 
  geom_point(alpha = .75)
```

## Add groups 

`group_by()` makes these groupings explicit so that they can be included in subsequent operations. For example, we might group weather_df by name and month:

```{r}
weather_df |>
  group_by(name, month)
```

grouping and counting things 
(grouping by name and summarizing what u have)
```{r}
weather_df |>
  group_by(name,month) |>
  summarize(
    n =n()
  )
```

grouping with an n_distinct 
 (will count u distinct instances not just number)
 
```{r}
weather_df |>
  group_by(month) |>
  summarize(
    n = n_distinct(date))
```
 
you can count directly

```{r}
weather_df |>
  count(name)
```

## More interesting summaries 

```{r}
weather_df |>
  group_by(name, month) |>
  summarize(
    mean_tmax = mean(tmax, na.rm =TRUE),
    median_tmin = median(tmin, na.rm =TRUE),
    sd_prcp =sd(prcp, na.rm = TRUE)
  )
```



```{r}
weather_df |>
  group_by(name,month) |>
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) |>
  ggplot(aes( x= month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_line()
```


can see what things look like seasonally and make it more visually appealing with pivot_wider

```{r}
weather_df |>
  group_by(name,month) |>
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) |>
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) |>
  knitr::kable(digits = 2)
```


## mutate with groups 

```{r}
weather_df |>
  group_by(name) |>
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    center_tmax = tmax - mean_tmax
  ) |> 
  ggplot(aes( x = date, y = center_tmax, color = name)) +
  geom_point()
```


Looking for cold days 

```{r}
weather_df |>
  group_by(name, month) |>
  mutate(temp_rank = min_rank(desc(tmin))) |>
  filter(temp_rank < 2)
```

What about lags?


```{r}
weather_df |>
  group_by(name) |>
  mutate(lagged_tmax = lag(tmax))
```

use the variables you create

```{r}
weather_df |>
  group_by(name) |>
  mutate(
    temp_change = tmax - lag(tmax)
) |> 
  summarize(
    sd_tmax_change = sd(temp_change, na.rm = TRUE),
    tmax_change_max = max(temp_change, na.rm = TRUE)
  )
```


```{r}
weather_df |>
  group_by(name) |>
  mutate(
    temp_change = tmax - lag(tmax),
    change_rank = min_rank(desc(temp_change))
) |> 
  filter(
    change_rank < 2
  )
```

Leanring Assessment: write a code chunk that imports, cleans, and summarizes the PULSE data to examine the mean and median at each visit. Export the results of this in a reader-friendly format.

```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") |>
  select(id, visit, everything()) |>
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) |>
  arrange(id, visit)

pulse_df |> 
  group_by(visit) |> 
  summarize(
    mean_bdi = mean(bdi, na.rm = TRUE),
    median_bdi = median(bdi, na.rm = TRUE)) |> 
  knitr::kable(digits = 3)
```









