---
title: "Viz 9/25"
author: "Kennedy Wade"
date: "2025-09-25"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
library(p8105.datasets)
```

Import the dataset

```{r}
data("weather_df")
```

```{r}
ggplot(data = weather_df, mapping = aes(x=tmin, y=tmax)) +
  geom_point()
```

```{r}
waether_df |>
  ggp;lot(aes(x=tmin, y=tmax)) + 
  geom_point()

ggp_weather_scatterplot
```

check that some rows are missing

```{r}
weather_df |>
  filter(is.na(tmax))
```

##fancier Scatterplots!
Different aestheics functions:
alpha makes the ppints on the scatterplot more transparent to see the density
geom_smooth creates tge trendline
all the lines that go under the ggplot are applied unless specified otherwise 
prcp is precipitation variable in the climate dataset 
```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax, color = name)) +
  geom_point(aes(color = name, size =prcp), alpha = 0.3) +
  geom_smooth(se =FALSE)
```

use faceting 

```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax, color = name)) +
  geom_point(aes(color = name, size =prcp), alpha = 0.3) +
  geom_smooth(se =FALSE) +
  facet_grid(.~name)
```

```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax, color = name)) +
  geom_point(aes(color = name, size =prcp), alpha = 0.3) +
  geom_smooth(se =FALSE) +
  facet_grid(name ~.)
```

lets make a more interesting scatterplot (shape)

```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax, color = name, shape = name)) +
  geom_point(aes(color = name, size =prcp), alpha = 0.3) +
  geom_smooth(se =FALSE) +
  facet_grid(.~name)

```

Learning assessment plot 

```{r}
weather_df |> 
filter(name == "CentralPark_NY") |> 
mutate(
  tmax_fahr = tmax * (9/5) +32,
  tmin_fahr = tmin * (9/5) +32
 ) |>
 ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
 geom_point() +
 geom_smooth(se = FALSE, method = "lm")
```

##small things 
you can take things out by commenting them out (add#) 
geom_smoooth should be last funtion
```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax, color = name, shape = name)) +
 # geom_point(alpha = 0.3) +
  geom_smooth(se =FALSE) 

weather_df |>
  ggplot(aes(x=tmin, y=tmax, color = name, shape = name)) +
  geom_smooth(se =FALSE) 
  geom_point()

```

adding hexagonal plot 

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax)) +
  geom_hex()
  
```


```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(color = "purple")
```

##univariate point
making a historgram 
color is for the outline, fill is for the inside of the bars 
```{r}
weather_df |> 
  ggplot(aes(x = tmin)) +
  geom_histogram(color = "white", fill = "red")
```

overlapping histogram

```{r}
weather_df |> 
  ggplot(aes(x = tmin, color = name)) + 
  geom_histogram()

weather_df |> 
  ggplot(aes(x = tmin, fill = name)) + 
  geom_histogram()
```

add facet grid to make sure they are not overlapping

```{r}
weather_df |> 
  ggplot(aes(x = tmin, fill = name)) + 
  geom_histogram() +
  facet_grid(name~.)
```

or you coudld make a desnsity plot 

```{r}
weather_df |> 
  ggplot(aes(x = tmin, fill = name)) + 
  geom_density() 
```

making a boxplot 

```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmin)) + 
  geom_boxplot(aes(fill = name))
```

violin plot 

```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmin, fill = name)) + 
  geom_violin()
```

ridge plot

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = name, fill = name)) + 
  geom_density_ridges()
```

learning assessment univariate plots 

learning assessment: make plots that com,pare precipitation across locations. try histogram, a density plot,  boxplot, a violin plot, and a ridgeplot; Uuse asethetic mappings to make your figure readable 

```{r}
weather_df |> 
  ggplot(aes(x =prcp, fill = name)) + 
  geom_density(alpha = 0.2)

weather_df |> 
  ggplot(aes(x =prcp, fill = name)) + 
  geom_histogram() +
  facet_grid(name~.)

weather_df |> 
  filter(prcp >5, prcp < 1000) |>
  ggplot(aes(x =prcp, fill = name)) +
  geom_density(alpha = 0.2)

```

#saving plots 

```{r}
ggp_weather_violin = 
  weather_df |> 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin()

ggp_weather_violin

ggsave("plots/violin_plot.pdf", ggp_weather_violin,
       width = 8, height = 6)
```


embedding plots 

```{r, fig.width = 6, fig.height =4}
  ggp_weather_violin
```










































