---
title: "visualizations 2 9.30.25"
author: "Kennedy Wade"
date: "2025-09-30"
output: github_document
---

```{r}
library(tidyverse)
library(patchwork)
library(p8105.datasets)
library(viridis)
data("weather_df")
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5)
```

lets make a basic scatterplot
##labs add labels to the graphs


```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from NOAA"
  )
```


##scales (update my base plot )
can chnage scale x or y axis limits, zooms into only a section of the plot that you care about  scale_color_hue o=changes color scheme but jeff wats us to use viridis bc its friendlier 

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15º C", "0", "15")
  )+
  scale_y_continuous(
    trans = "sqrt", 
    position = "right")

 viridis::scale_color_viridis(
    discrete = TRUE
  )

```

## save a big code chunk as a name called "ggp_temperature so u dont have to type it all out. could also use ggsave to the scatterplot 



```{r}
ggp_temperature = 
 weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package") + 
   scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15º C", "0", "15")
  )+
  scale_y_continuous(
    trans = "sqrt", 
    limits = c(10,30)
  )+
 viridis::scale_color_viridis(
    discrete = TRUE
  )

```


```{r}
ggp_temperature =
  theme_minimal()+
  theme(legend.position = "bottom")
 
```

##adding data in geoms

```{r}
central_park_df = 
  weather_df |> 
  filter(name == "CentralPark_NY")

molokai_df = 
  weather_df |> 
  filter(name == "Molokai_HI")

ggplot(data = molokai_df, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park_df) 
```

##patchwork!!: making 3 plots and combining them using patchwork to squeeze mutiple plots in th esame panel to have a lot of info without taking p alot of space 

```{r}
ggp_tmax_tmin = 
  weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

ggp_prec_density = 
  weather_df |> 
  filter(prcp > 0) |> 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

ggp_temp_season = 
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(ggp_tmax_tmin + ggp_prec_density) / ggp_temp_season
```


## Data Manipulation

The behavior of your plot depends on the data you’ve supplied; in some cases, it’s easier to control behavior through data manipulation than it is through the plot code. You can change the order level of a factor variable to your specified preference using forcats::fct_relevel or according to the value of another variable using forcats::fct_reorder.


making more temperature violin plots. factor rlevel assigns molokai, cp, and waterhole in the order u want 
```{r}
weather_df |>
  mutate(name = forcats::fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) |> 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

second exaple reoders name according to tmax values in each name 
```{r}
weather_df |>
  mutate(name = forcats::fct_reorder(name, tmax)) |> 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```


```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") |>
  select(id, visit, everything()) |>
  mutate(
    visit = recode(visit, "bl" = "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) |>
  arrange(id, visit)

ggplot(pulse_data, aes(x = visit, y = bdi)) + 
  geom_boxplot()
```

make a plot for the FAS study 

```{r}
pups_df = 
 read_csv("./data/FAS_pups.csv", na = c("NA", ".", ""), skip = 3) |>
  janitor::clean_names() |>
  mutate(
    sex = 
      case_match(
        sex, 
        1 ~ "male", 
        2 ~ "female"))

litters_df = 
  read_csv("./data/FAS_litters.csv", na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_df = left_join(pups_df, litters_df, by = "litter_number") 

fas_df |> 
  select(sex, dose, day_of_tx, pd_ears:pd_walk) |> 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day") |> 
  drop_na() |> 
  mutate(outcome = forcats::fct_reorder(outcome, pn_day, median)) |> 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() + 
  facet_grid(day_of_tx ~ outcome)
```




















