---
title: "Viz 9/25"
author: "Kennedy Wade"
date: "2025-09-25"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
library(p8105.datasets)
```

Import the dataset

```{r}
data("weather_df")
```

```{r}
ggplot(data = weather_df, mapping = aes(x=tmin, y=tmax)) +
  geom_point()
```

```{r}
waether_df |>
  ggp;lot(aes(x=tmin, y=tmax)) + 
  geom_point()

ggp_weather_scatterplot
```

check that some rows are missing

```{r}
weather_df |>
  filter(is.na(tmax))
```

##fancier Scatterplots!
Different aestheics functions:
alpha makes the ppints on the scatterplot more transparent to see the density
geom_smooth creates tge trendline
all the lines that go under the ggplot are applied unless specified otherwise 
prcp is precipitation variable in the climate dataset 
```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax, color = name)) +
  geom_point(aes(color = name, size =prcp), alpha = 0.3) +
  geom_smooth(se =FALSE)
```

use faceting 

```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax, color = name)) +
  geom_point(aes(color = name, size =prcp), alpha = 0.3) +
  geom_smooth(se =FALSE) +
  facet_grid(.~name)
```

```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax, color = name)) +
  geom_point(aes(color = name, size =prcp), alpha = 0.3) +
  geom_smooth(se =FALSE) +
  facet_grid(name ~.)
```

lets make a more interesting scatterplot (shape)

```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax, color = name, shape = name)) +
  geom_point(aes(color = name, size =prcp), alpha = 0.3) +
  geom_smooth(se =FALSE) +
  facet_grid(.~name)
```
```
Learning assessment plot 

```{r}

```















